---
title: "Problem Set 1"
author: "Sachin Badole"
date: "9/17/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r include=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)

library(data.table)
library(rlang)
suppressMessages(library(dplyr))  # I am using the "dplyr" package.
library(ggplot2)

```

Economics 725: Machine Learning for Economists, University of Wisconsin-Madison. \

Question (1)\
0) I have download Rstudio. \
1) I have download the DB1BMarket data table as per given instrucation in problem set 1. \
 a. Take only data from the first quarter of 2015.- Done \
 b. Take the following variables: ItinID, MktID, OriginAirportID, DestAirportID, TkCarrierChange,        TicketCarrier, Passengers, MarketFare, and MarketDistance. - Done \
 c. Download the data and bring it into R. - Done

```{r} 
# Set the working directory.
setwd("G:/My Documents/Sem III/Econ 725 Machine Learning for Econmist/Problem Sets/Problem set 1")

# Load dataset using the following command.
Airline_ticktes_data <- read.csv("65799243_T_DB1B_MARKET.csv")

# 1) b) Here, I am checking the name of the columns.
names(Airline_ticktes_data)    # Name of columns

```

```{r}
# I found there is one extra column in the above dataset, so I have dropped it.
Airline_ticktes_data <- subset(Airline_ticktes_data, select = -c(X) )  
```
Initial number of observations.
```{r}
length(Airline_ticktes_data$ITIN_ID)  # Initial number of observations.
```

Question (2) \
Remove tickets that can't be assigned to a unique carrier, remove markets (a unidirectional origin-destination pair) with less than 20 passengers per day, and remove tickets with extreme prices.\

```{r}
# 2.a) Removing any tickets taht have a ticket carrier change. Ticketing Carrier Change Indicator (1=Yes)
Airline_ticktes_data_v1 <- Airline_ticktes_data[Airline_ticktes_data$TK_CARRIER_CHANGE==0,]
```

The number of observation after Removing any tickets taht have a ticket carrier change.
```{r}
# shows number of observatins after the removing the ticketing carrier.
length(Airline_ticktes_data_v1$TK_CARRIER_CHANGE) 
```

```{r}
# 2.c) Remove tickets with prices less than $25 or more than $2,500. 
Airline_ticktes_data_v1 <- Airline_ticktes_data_v1[Airline_ticktes_data_v1$MARKET_FARE>25 & Airline_ticktes_data_v1$MARKET_FARE<2500,]
```

The number of observation after remove tickets with prices less than $25 or more then $2500.
```{r}
# The number of observation after remove tickets with prices less than $25 or more then $2500.
length(Airline_ticktes_data_v1$MARKET_FARE)
```

```{r}
# 2.b ) Create new Variable called Total_No_Passengers which defind that passengers number multiply by 10 for each ticket.

# Find the total numebr of passengers in each market.
Airline_ticktes_data_v1 <- Airline_ticktes_data_v1 %>%
  group_by(ORIGIN_AIRPORT_ID, DEST_AIRPORT_ID) %>%
  mutate(TOTAL_PASSENGERR = sum(PASSENGERS)<(365/4)*20/10)

# Drop some observations those are duplicates.
Airline_ticktes_data_v1 <- Airline_ticktes_data_v1[!(Airline_ticktes_data_v1$TOTAL_PASSENGERR==TRUE),]
```

The number of observation after find the total number of passengers in each market.
```{r}
# The number of observation after find the total number of passengers in each market.
length(Airline_ticktes_data_v1$TOTAL_PASSENGERR)

```

Question (3)- You will create two datasets: one at the market-carrier level and another at the market level.

a) For each market-airline. (Calculate the average price, Calculate the total number of passengers, and Calculate the average distance.)\

```{r}
# a) For each market-airline.

data_market_airline <- Airline_ticktes_data_v1 %>%
  group_by(ORIGIN_AIRPORT_ID, DEST_AIRPORT_ID, TICKET_CARRIER) %>%
  mutate(TOTAL_NO_PASSENGERS = sum(PASSENGERS)*10, 
         AVERAGE_PRICE_AIRLINE = weighted.mean(MARKET_FARE, PASSENGERS),
         AVERAGE_DISTANCE_AIRLINE = weighted.mean(MARKET_DISTANCE,PASSENGERS))

data_market_airline <- data_market_airline %>% 
  group_by(ORIGIN_AIRPORT_ID, DEST_AIRPORT_ID, TICKET_CARRIER) %>%
  distinct(TOTAL_NO_PASSENGERS, AVERAGE_PRICE_AIRLINE, AVERAGE_DISTANCE_AIRLINE)
```

The number of observation in the Market-airline dataset.
```{r}
# The number of observation in the Market-airline dataset.
length(data_market_airline$TOTAL_NO_PASSENGERS)
```

```{r}
# b)For each market

data_market <- data_market_airline %>%
  group_by(ORIGIN_AIRPORT_ID, DEST_AIRPORT_ID) %>%
  mutate(
         AVERAGE_PRICE_MARKET = weighted.mean(AVERAGE_PRICE_AIRLINE, TOTAL_NO_PASSENGERS) ,                             AVERAGE_DISTANCE_MARKET = weighted.mean(AVERAGE_DISTANCE_AIRLINE, TOTAL_NO_PASSENGERS),
         HHI = sum(((TOTAL_NO_PASSENGERS*100)/sum(TOTAL_NO_PASSENGERS))^2))   

data_market <- data_market %>%
              count(DEST_AIRPORT_ID,AVERAGE_PRICE_MARKET,AVERAGE_DISTANCE_MARKET, HHI, sort = TRUE, name = "TOTAL_NO_FIRMS")

colnames(data_market) = tolower(colnames(data_market))
```

The number of observation in the Market-level dataset.
```{r}
# The number of observation in the Market-level dataset.
length(data_market$hhi)
```

Load the given populations data and merge with the market-level dataset.\
```{r}
# load given populations data for merge.
load(file="populations.R")

#data_market <- merge(data_market,populations,by.x="origin_airport_id", by.y="dest_airport_id")
data_market <- merge(data_market,populations,by=c("origin_airport_id","dest_airport_id"))
```

The number of observation in the Market-level dataset after merging with population data.
```{r}

# The number of observation in the Market-level dataset after merging with population data.
length(data_market$hhi)
```

4) Generate tables with summary statistics for each of your datasets and generate plots characterizing the distributions of market level prices and HHI as well as the relationship between them.\

a) Report summary statistics for your tables (hint: use the kable function in the knitr package).\

Summary Statistics for each market-airline.\

```{r}

kable(summary(data_market_airline[,c("TOTAL_NO_PASSENGERS","AVERAGE_PRICE_AIRLINE","AVERAGE_DISTANCE_AIRLINE")]))

```

Summary Statistics for each market.\

```{r}
View(data_market)
kable(summary(data_market[,c("average_price_market", "average_distance_market", "hhi", "market_size", "total_no_firms")]))
```

b) Plots\

1. The Scatter plot of HHI versus prices at the market level.\

```{r}
ggplot(data_market,aes(x=hhi,y=average_price_market)) + 
  geom_point(color="blue") + theme_bw() + 
  labs( title = "Airline Price and Market Structure", x = "HHI", y = "Average Price") +
  theme(plot.title = element_text(hjust = 0.5)) 
```  

2. The market level HHI density plot.\

```{r}
 ggplot(data_market,aes(x=hhi)) + 
   labs( title = "HHI Density", x = "HHI", y = "Density")+
   geom_density(color="darkblue", fill="lightblue") + theme_bw() +
   theme(plot.title = element_text(hjust = 0.5))
```

3. The market level Average price density plot.\

```{r} 
ggplot(data_market,aes(x=average_price_market)) + 
   labs( title = "Price Density", x = "Price", y = "Density")+
   geom_density(color="darkblue", fill="lightblue") + theme_bw() +
   theme(plot.title = element_text(hjust = 0.5))
```












